Efficient Thresholded Correlation using Truncated Singular Value Decomposition

### 2015-07-22 threshold/draft.Rnw
Added TAPER reference

XXX add discussion of applications (genomics?) and previous work (in particular
\cite{latj} and \cite{bhpt} are related but different),
then on to the preliminaries:
\[\]

@@

In practice, many pairs of vectors may be ruled out in this fashion by
inspecting distances in low-dimensional subspaces, substantially reducing the
complexity of computing thresholded correlation matrices.  Approaches that
cheaply omit pairs of vectors from consideration are called \emph{pruning
methods}, and a variety of pruning methods have been described elsewhere for
the thresholded correlation problem; see, for example\cite{zf,mnl} and the
references therein.  The approach described here prunes along an intuitive path
of decreasing variance defined by the data and computed by the SVD.

@@

\begin{thebibliography}{99}
\bibitem{br}
Baglama, J. and Reichel, L., Augmented Implicitly Restarted Lanczos Bidiagonalization Methods, SIAM J. Sci. Comput., 2005.
\bibitem{bhpt}
Bair, E., Hastie, T., Paul, D.., and Tibshirani, P., Prediction by supervised principal components, Journal of the American Statistical Association 101, no. 473, 2006.
\bibitem{latj}
Levin, Asriel U., Todd K. Leen, and John E. Moody. Fast pruning using principal components. (1993).
\bibitem{mnl}
Mueen, A., Nath, S., and Liu, J., Fast approximate correlation for
massive time-series data, in Proceedings of the 2010 ACM SIGMOD International
Conference on Management of data, 2010, pp. 171-182.
\bibitem{raf}
Rafiei, D., On similarity-based queries for time series data,
in Data Engineering Proceedings., IEEE, 1999.
\bibitem{zf}
Zhang, J. and Feigenbaum, J, Finding Highly Correlated Pairs Efficiently with Powerful Pruning,
in Proceedings of ACM Conference on Information and Knowledge Management (CIKM), 2006, pp. 152-161.
\bibitem{irlba-R}
\url{https://cran.r-project.org/web/packages/irlba/index.html}
\bibitem{irlba-P}
\url{https://github.com/bwlewis/irlbpy}
\bibitem{irlba-M}
\url{http://www.math.uri.edu/~jbaglama/software/irlba.m}
\end{thebibliography}
\end{document}

>>>>

XXX add discussion of applications (genomics?) and previous work (especially
\cite{xstk1},\cite{xstk2} and
related but different work that use PCA: \cite{latj} and \cite{bhpt} are related but different),
then on to the preliminaries:
\[\]

@@

In practice, many pairs of vectors may be ruled out in this fashion by
inspecting distances in low-dimensional subspaces, substantially reducing the
complexity of computing thresholded correlation matrices.  Approaches that
cheaply omit pairs of vectors from consideration are called \emph{pruning
methods}, and a variety of pruning methods have been described elsewhere for
the thresholded correlation problem; see, for example\cite{xstk1,xstk2,zf,mnl}
and the references therein.  The approach described here prunes along an
intuitive path of decreasing variance defined by the data and computed by the
SVD.

@@

\begin{thebibliography}{99}
\bibitem{br}
Baglama, J. and Reichel, L., Augmented Implicitly Restarted Lanczos Bidiagonalization Methods, SIAM J. Sci. Comput., 2005.
\bibitem{bhpt}
Bair, E., Hastie, T., Paul, D.., and Tibshirani, P., Prediction by supervised principal components, Journal of the American Statistical Association 101, no. 473, 2006.
\bibitem{latj}
Levin, Asriel U., Todd K. Leen, and John E. Moody. Fast pruning using principal components. (1993).
\bibitem{mnl}
Mueen, A., Nath, S., and Liu, J., Fast approximate correlation for
massive time-series data, in Proceedings of the 2010 ACM SIGMOD International
Conference on Management of data, 2010, pp. 171-182.
\bibitem{xstk1}
Xiong, Hui, et al. "Exploiting a support-based upper bound of Pearson's correlation coefficient for efficiently identifying strongly correlated pairs." Proceedings of the tenth ACM SIGKDD international conference on Knowledge discovery and data mining. ACM, 2004.
\bibitem{xstk2}
Xiong, Hui, et al. "TAPER: A two-step approach for all-strong-pairs correlation query in large databases." Knowledge and Data Engineering, IEEE Transactions on 18.4 (2006): 493-508.
\bibitem{zf}
Zhang, J. and Feigenbaum, J, Finding Highly Correlated Pairs Efficiently with Powerful Pruning,
in Proceedings of ACM Conference on Information and Knowledge Management (CIKM), 2006, pp. 152-161.
\bibitem{irlba-R}
\url{https://cran.r-project.org/web/packages/irlba/index.html}
\bibitem{irlba-P}
\url{https://github.com/bwlewis/irlbpy}
\bibitem{irlba-M}
\url{http://www.math.uri.edu/~jbaglama/software/irlba.m}
\end{thebibliography}
\end{document}
